*Prochlorococcus marinus* responses to light and oxygen

### 2024-05-31 Output/PLOSRound2/ResponsetoReviewers.txt

The abstract seems to omit any discussion of the conclusions/bigger picture of the study.

-Thank you for comment. We have added text to discuss the main conclusions of this study.

Line 55: LLII/III and LLIV are not the ?oldest? clades, but the most basal or early branching.

-We have edited the text in the revised manuscript: "Clades LLII and LLIII, including cultured strain SS120, are
grouped together as early branching phylogenetic lineage in
the P. marinus radiation, with a preference for low light."
Line: XXX

@@

Line 369: The authors invoke ?limitations in the annotation process? to explain the lack of evidence for 
MED4 ATP synthase. What does this mean? The MED4 genome surely contains ATP synthase genes.

-We agree that the MED4 genome includes genes encoding ATP Synthase. Upon verifying our annotation assignments we discovered the problem with the absence of ATP Synthase from Clade HLI was not due to the limitations of the process but rather a mislabel of the column in figure 2. 'ATP Synthase' should have been labelled 'RUBISCO'
and 'RUBISCO' should have been labelled 'ATP Synthase'. However, in the Ocean Protein Portal database no RUBISCO peptides are annotated as coming from MED4. This likely reflects an annotation assignment issue because the RUBISCO proteins are highly conserved, so assignments to strains or clades of Prochlorococcus may be ambiguous.
We have edited the text in the revised manuscript to reflect the limitation of the annotation process as it pertains to the highly conserved RUBISCO complex: "Though present in the genome of MED4, the absence of proteins annotated as RUBISCO complex for Clade HLI, compared to annotated detections of RUBISCO across the other three clades suggests limitations in the OPP annotation process for highly conserved protein sequences."
Line: XXX

@@

Paragraph starting L363: Why is there no discussion of the Rubisco data? Or especially the absence 
thereof at low O2 for ecotypes HL1 and LL1? Since Rubisco is one of the most abundant proteins in a 
photosynthesizing cell, does this indicate that these ecotypes are present but not actively fixing carbon 
under these conditions? Also, is it worth noting that LLIV is the only variety to be detected at all in the 
deepest low O2 site -- which would appear to be consistent with subsequent findings in this study?

- We revised the text from lines 668 to XXX to incorporate these helpful observations. The following was added: " Notwithstanding limitations on assignments of proteins to clades, detections of the abundant carbon fixation complex RUBISCO derived from HLI, LLI and LLII/III were notably absent from samples taken at lower [O~2~], suggesting limited capacity for carbon fixation by HLI,  LLI and LLII/III under low [O~2~] habitats."

@@

- Thank you for this suggestion. We have combined the growth rate figure with the GAM model for each strain to faciliate comparisons.

I'm also not sure I understand why the red light data is relevant -- perhaps by moving it to supplemental, it would be easier to compile these data into compound figures?

We briefly motivated the inclusion of the red light growth trials, as mechanistically informative, at line 706:
"As mentioned, growth trials under red light, although not representative of *Prochlorococcus* niches, are mechanistically informative [@murphyPhotoinactivationPhotosystemII2017] regarding photoinactivation of PSII."

-Through the Discussion, notably under the discussion of the GAM models and responses to growth rate to increasing PUR, we return to the responses to red light to show that some 'high light' limitations on growth are interpretable as photoinhibition of PSII, which is more severe under blue, compared to red, light, notably at lines 1023-1032 (xXXX).

### 2024-05-31 Code/Manuscript_O2.Rmd
minor edits

Proteins from 13 annotated strains of *P. marinus* were detected across
depths and oxygen concentrations in the OPP data set
analyzed. We focused our analysis here on core photosynthetic protein
complexes, for clades HLI, LLI, LLII/III and LLIV (Fig \@ref(fig:CladeProchloroPhotosynthDepthO2)) as a function of depth (a
proxy for light intensity) and measured [O~2~] at the sampling locations. Photosynthetic complexes
from HLI (including strain MED4) were detected throughout the water
column, but, with one exception, only under high [O~2~]. Though present in the genome of MED4, the absence of proteins annotated as ATP Synthase complex for Clade HLI, compared to annotated
detections of ATP Synthase across the other three clades, suggests
limitations in the annotation process for highly conserved protein
sequences. Complexes from Clade LLI (including strain NATL2A) were also
present across the depth/light axes, although with more protein detections at lower
[O~2~], and fewer near surface detections of proteins from Clade LLI. Notwithstanding limitations on assignments of proteins to clades, detections of the abundant carbon fixation complex RUBISCO derived from HLI and LLI were notably absent from samples taken at lower [O~2~], suggesting limited capacity for carbon fixation by HLI and LLI under low [O~2~] habitats. Complexes from clade LLII/III
(including strain SS120) were also detected across the depth/light and [O~2~] ranges,
with more detections at deeper, darker depths. Clade LLIV (including strain MIT9313)
photosynthetic complexes were also detected throughout the depth/light and [O~2~], ranges, notably with the most frequent detections of proteins at both depth and at low [O~2~], compared to other strains. Our analyses utilized a proteomics dataset; however, alternate approaches, such as metagenomics or metatranscriptomics, could have been employed to analyze ecotype abundances using the TARA Oceans [@pesantOpenScienceResources2015] or other datasets.

@@

# This is Fig2.tiff
CladeProchloroPhotosynthDepthO2_cap <- glue("**Ocean detection of *Prochlorococcus marinus* photosynthesis complexes.**", 
                       "Protein detections (circles) are plotted vs. O~2~ (µM) (X-axis) and depth (m) (Y-axis) at sample origin.",
                       "Rows separate data annotated as from *Prochlorococcus marinus* clades: HLI (including *P. marinus* MED4, solid black circles), LLI (including *P. marinus* NATL2A, solid black circles), LLII/III (including *P. marinus* SS120, solid black circles) and LLIV (including *P. marinus* MIT9313, solid black circles).",
                       "Columns show detections of proteins annotated as  Photosystem II (PSII), Cytochromeb6f complex (Cytb6f), Photosystem I (PSI), ATP Synthase or Ribulose-1,5-bisphosphate oxygenase carboxylase (RUBISCO).",
                       "For comparison, culture growth rate experimental conditions are indicated by horizontal grey lines for depths approximating peak Photosynthetically Active Radiation (PAR; µmol photons m^-2^ s^-1^); and vertical grey lines for imposed [O~2~] (µM).",
                       "Data obtained from the Biological and Chemical Oceanography Data Management Office repository [@saitoPeptidesTheirSpectral2018].",
 .sep = " ")

@@

```{r CladeProchloroProteinMetabDepthO2, fig.cap = CladeProchloroProteinMetabDepthO2_cap , fig.height = 20, fig.width = 10, echo = FALSE}

# This is Fig7.tiff
CladeProchloroProteinMetabDepthO2_cap <- glue("**Ocean detection of *Prochlorococcus marinus* protein metabolism complexes.**", 
                       "Protein detections (circles) are plotted vs. O~2~ (µM) (X-axis) and depth (m) (Y-axis) at sample origin.",
                       "Rows separate data annotated as from *Prochlorococcus marinus* clades: HLI (including *P. marinus* MED4, solid black circles), LLI (including *P. marinus* NATL2A, solid black circles), LLII/III (including *P. marinus* SS120, solid black circles) and LLIV (including *P. marinus* MIT9313, solid black circles).",
                       "Columns show detections of proteins annotated as FtsH Protease Complexes (FtsH1, FtsH2, FtsH3) or the Ribosome.",
                       "For comparison, culture growth rate experimental conditions are indicated by horizontal grey lines for depths approximating peak Photosynthetically Active Radiation (PAR; µmol photons m^-2^ s^-1^); and vertical grey lines for [O~2~] (µM).",
                       "Data obtained from the Biological and Chemical Oceanography Data Management Office repository [@saitoPeptidesTheirSpectral2018].",
 .sep = " ")

@@

```{r NichesMap, fig.cap = NichesMap_cap, fig.height = 15, fig.width = 15, echo = FALSE}

@@

knitr::include_graphics(file.path(FiguresIn, "MapComboPro.png"))

>>>>

Proteins from 8 annotated strains of *P. marinus* were detected across
depths and oxygen concentrations in the OPP data set
analyzed. We focused our analysis here on core photosynthetic protein
complexes, for clades HLI (2 strains), LLI (2 strains), LLII/III (2 strains) and LLIV (2 strains) (Fig \@ref(fig:CladeProchloroPhotosynthDepthO2)) as a function of depth (a
proxy for light intensity) and measured [O~2~] at the sampling locations. Photosynthetic complexes
from HLI (including strain MED4) and LLI (including strain NATL2A) were detected throughout the water
column, predominately under high [O~2~]. Though present in the genome of MED4, the absence of proteins annotated as RUBISCO complex for clade HLI, compared to annotated
detections of RUBISCO across the other three clades, suggests
limitations in the annotation process for highly conserved protein
sequences. Notwithstanding limitations on assignments of proteins to clades, detections of the abundant carbon fixation complex RUBISCO derived from HLI, LLI and LLII/III were notably absent from samples taken at lower [O~2~], suggesting limited capacity for carbon fixation by HLI,  LLI and LLII/III under low [O~2~] habitats. Clade LLII/III (including strain SS120)
photosynthetic complexes were also detected throughout the depth/light and [O~2~] ranges, notably with the most frequent detections of proteins at both high and low [O~2~], compared to other strains. Complexes from LLIV (including strain MIT9313) were also detected across the depth/light and [O~2~] ranges, with more detections at deeper, darker depths and at low [O~2~].Our analyses utilized a proteomics dataset; however, alternate approaches, such as metagenomics or metatranscriptomics, could have been employed to analyze ecotype abundances using the TARA Oceans [@pesantOpenScienceResources2015] or other datasets.

@@

# This is Fig2.tiff
CladeProchloroPhotosynthDepthO2_cap <- glue("**Ocean detection of *Prochlorococcus marinus* photosynthesis complexes.**", 
                       "Protein detections (circles) are plotted vs. O~2~ (µM) (x-axis) and depth (m) (y-axis) at sample origin with a 15% offset to seperate protein detections occupying the same origin.",
                       "Rows separate data annotated as from *Prochlorococcus marinus* clades: HLI (including *P. marinus* MED4, solid black circles), LLI (including *P. marinus* NATL2A, solid black circles), LLII/III (including *P. marinus* SS120, solid black circles) and LLIV (including *P. marinus* MIT9313, solid black circles).",
                       "Columns show detections of proteins annotated as Photosystem II (PSII), Cytochromeb6f complex (Cytb6f), Photosystem I (PSI), ATP Synthase or Ribulose-1,5-bisphosphate oxygenase carboxylase (RUBISCO).",
                       "For comparison, culture growth rate experimental conditions are indicated by horizontal grey lines for depths approximating peak Photosynthetically Active Radiation (PAR; µmol photons m^-2^ s^-1^); and vertical grey lines for imposed [O~2~] (µM).",
                       "Data obtained from the Biological and Chemical Oceanography Data Management Office repository [@saitoPeptidesTheirSpectral2018].",
 .sep = " ")

@@

```{r CladeProchloroProteinMetabDepthO2, fig.cap = CladeProchloroProteinMetabDepthO2_cap , fig.height = 20, fig.width = 20, echo = FALSE}

# This is Fig7.tiff
CladeProchloroProteinMetabDepthO2_cap <- glue("**Ocean detection of *Prochlorococcus marinus* protein metabolism complexes.**", 
                       "Protein detections (circles) are plotted vs. O~2~ (µM) (x-axis) and depth (m) (y-axis) at sample origin with a 15% offset to seperate protein detections occupying the same origin.",
                       "Rows separate data annotated as from *Prochlorococcus marinus* clades: HLI (including *P. marinus* MED4, solid black circles), LLI (including *P. marinus* NATL2A, solid black circles), LLII/III (including *P. marinus* SS120, solid black circles) and LLIV (including *P. marinus* MIT9313, solid black circles).",
                       "Columns show detections of proteins annotated as FtsH Protease Complexes (FtsH1, FtsH2, FtsH3) or the Ribosome.",
                       "For comparison, culture growth rate experimental conditions are indicated by horizontal grey lines for depths approximating peak Photosynthetically Active Radiation (PAR; µmol photons m^-2^ s^-1^); and vertical grey lines for [O~2~] (µM).",
                       "Data obtained from the Biological and Chemical Oceanography Data Management Office repository [@saitoPeptidesTheirSpectral2018].",
 .sep = " ")

@@

```{r NichesMap, fig.cap = NichesMap_cap, fig.height = 20, fig.width = 20, echo = FALSE}

@@

knitr::include_graphics(file.path(FiguresIn, "NichesMap.png"))
